\lesson{1}{10/28/2022}{Introduction}

\begin{itemize}
    \item First part: fundamental groups. We follow Munkres:
        \begin{itemize}
            \item Chap 9 `Fundamental group'
            \item Chap 11 `The Seifert-Van Kampen thm' and
            \item Chap 12/13 `Classification covering spaces'
        \end{itemize}
    \item Second part: Homology groups, via cudi, chapter of a book.
    \item 10 problems (solve them during the semester) No feedback during the semester, but asking questions is allowed. Working together is allowed.
    \item Exam (completely open book) First 4 questions: 1h30 prep. Last one 30 min, no prep
        \begin{enumerate}
            \item Theoretical question (open book, explain the proof, \ldots) \hfill /4
            \item New problem (comparable to one of the 10 problems)\hfill /4
            \item Explain your solution $n$-th problem solved at home \hfill /4
            \item Explain your solution $m$-th problem solved at home\hfill /4
            \item 4 small questions \hfill /4
        \end{enumerate}
        After the exam, hand in your solutions of the other problems. After a quick look, the points of 3.\ and 4.\ can be $\pm 1$ (in extreme cases $\pm 2$)
    \item No exercise classes for this course!
\end{itemize}

\setcounter{chapter}{0}
\chapter{Introduction}

\section{What is algebraic topology?}
Functor from category of topological spaces to the category of groups.
\begin{itemize}
    \item Category: set of spaces and morphisms.
    \item Functor: $X \leadsto G_X$ and  $f:X\to Y \leadsto f_*: G_X \to  G_Y$ such that
        \begin{itemize}
            \item $(f \circ g)_* = f_* \circ g_* $
            \item $(1_X)_* = 1_{G_X}$
        \end{itemize}
\end{itemize}

Two systems we'll discuss:
\begin{itemize}
    \item fundamental groups
    \item homology groups
\end{itemize}


\begin{eg}
    Suppose we have a functor.
    If $G_X \not\cong G_Y$, then  $X$ and  $Y$ are not homeomorphic.
    If `shadows' are different, then objects themselves are different too.
\end{eg}
\begin{explanation}
    Suppose $X$ and $Y$ are homeomorphic.
    Then $\exists f: X \to  Y$ and $g: Y \to  X$, maps (maps are always continuous in this course), such that $g  \circ f = 1_X$ and $f \circ g = 1_Y$.
    Then $f_*: G_X \to G_Y$ and $g_*: G_Y \to  G_X$ such that $(g \circ f)_* = (1_X)_*$ and  $(f \circ g)_* = (1_Y)_*$. Using the rules discussed previously, we get
    \[
    g_*  \circ f_* = 1_{G_X} \quad f_*  \circ  g_* = 1_{G_Y}
    ,\] 
    which means that $f_* : G_X \to  G_Y$ is an isomorphism.
\end{explanation}